---
title: "Guide to estimatr"
author: "Aakriti"
date: '2023-04-10'
output: html_document
---

# Overview

The estimatr package is a specialized tool in R that offers quick and efficient estimators for commonly used research designs in the social sciences. Several of the estimators available in the R or other popular packages are sluggish and can produce inappropriate estimates due to their default settings. Moreover, some of the latest and most advanced estimators have not yet been conveniently incorporated into R packages. `estimatr` aims to address these issues by offering estimators that are optimized for design-based inference, resulting in more accurate and reliable estimates.

Using the `wage1` dataset, we will walk you through the process of estimating regression coefficients while accounting for potential sources of bias and violations of statistical assumptions using two of the most commonly used functions in estimatr: `lm_robust()` and `iv_robust()`.

- explain the wage 1 dataset here: make a table of `variable | Description ` 

# Load Packages

```{r}

# Load the estimatr package
library(estimatr)
# Load the wage1 dataset from the wooldridge package
library(wooldridge) 
data("wage1")


```




# Estimate Linear Models using lm_robust

In regression analysis, standard errors are important for assessing the precision of estimates and determining whether the results obtained from a sample accurately represent the underlying population. In this example, we are interested in examining how the education level of employees relates to their wages. Let's first use the built-in lm() function to estimate this linear model. 

- write down the estimation equation here


```{r}

reg <- lm(data=wage1, wage~educ)
summary(reg)

#https://declaredesign.org/r/estimatr/articles/mathematical-notes.html for the explanations
```

- describe the homoscedasticity assumption briefly, why is it important. 
- add code and plots to check if the assumption holds 

However the lm()function is designed to fit linear models which assume homoscedasticity i.e. a constant error term along all independent variables. In the case of heteroscedasticity, robust standard errors are a way to get unbiased standard errors. 



This `lm_robust()` function is used to get the robust standard errors from a linear regression model. Now, let's re-estimate the model using robust standard errors. 

```{r}
lm_robust(data=wage1, wage~educ, se_type = "HC2")
```
The result shows that the standard errors are different from those obtained when homoscedasticity is assumed.

## Heteroskedasticity-Robust Standard Errors

The lm_robust() function in the estimatr package provides several options for calculating standard errors of regression coefficients using the se_type argument. The available options are:

se_type = "classical": This is the default option, which uses the classical or ordinary least squares (OLS) estimator to calculate standard errors. It assumes that the error terms are homoscedastic and uncorrelated with the independent variables.

se_type = "HC0": This option uses the heteroscedasticity-consistent (HC0) estimator to calculate standard errors. It allows for heteroscedasticity in the error terms, but does not correct for small sample bias.

se_type = "HC1": This option uses the HC1 estimator to calculate standard errors. It is similar to the HC0 estimator, but includes a small sample correction to improve the accuracy of standard errors.

se_type = "HC2": This option uses the HC2 estimator to calculate standard errors. It is similar to the HC1 estimator, but includes a different small sample correction.

se_type = "HC3": This option uses the HC3 estimator to calculate standard errors. It is similar to the HC2 estimator, but includes a more robust small sample correction that is less sensitive to outliers.

In summary, these options allow the user to choose between different types of standard errors that are appropriate for different assumptions about the error terms and sample size. The HC estimators are particularly useful when there is heteroscedasticity in the error terms, which violates the assumptions of the classical estimator.

## Cluster-Robust Standard Errors
Data is considered to be clustered when there are subsamples within the data that are related to each other. When error terms are correlated within clusters but independent across clusters, then regular standard errors, which assume independence between all observations, will be incorrect. Cluster-robust standard errors are designed to allow for correlation between observations within cluster. 

For cluster-robust inference, this package provides several estimators that are essentially analogs of the heteroskedastic-consistent variance estimators for the clustered case: `CR0` and `CR2`. The default is the `CR2` variance estimator, analogous to HC2 standard errors, and perform quite well in small samples without sacrificing much in the way of efficiency in larger samples.


# Tip:
Check out the [mathematical notes](https://declaredesign.org/r/estimatr/articles/mathematical-notes.html) for each of the estimators to better understand the formulas used to compute these standard errors and a discussion on the use-cases for the different standard errors and cluster types.

# Estimate Instrumental Variable (IV) regressions using iv_robust

- provide brief description on IV regressions, when is it useful 

- Explain the example case 

- specify the model/equation we estimate in this here

- specify with robust se 

- add other relevant content 



+ iv_robust 
  *2SLS estimation of IV regression
```{r}

# good IV = correlated with x but not with u; maybe experience 

iv_robust(data=wage2, wage~educ|IQ)



```

# Summary 

- write bullet point summaries of key take-aways from this building block