---
title: "Guide to estimatr"
author: "Aakriti"
date: '2023-04-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Guide To Estimatr

```{r}
install.packages("wooldridge")

install.packages("npsf")
library(npsf)

data(mroz)
data(hprice1)
data("nswre74")


```

Standard errors are essential for researchers running a regression as they assess how accurately the sample data represents the total population. The standard errors of a regression on R are obtained using the lm() function as so:
```{r}
library(wooldridge)
reg <- lm(data=wage1, wage~educ)
summary(reg)

#https://declaredesign.org/r/estimatr/articles/mathematical-notes.html for the explanations
```

However the lm()function is designed to fit linear models which assume homoscedasticity i.e. a constant error term along all independent variables. 

In the case of heteroscedasticity, robust standard errors are a way to get unbiased standard errors. 

R offers a package called 'Estimatr' which provides various functions to get estimates with their robust errors.

```{r}
install.packages("estimatr")
library(estimatr)

```


## Estimators provided by Estimatr


+ lm_robust  
  *

```{r}
lm_robust(data=wage1, wage~educ+exper)
#se_type = classical gives the lm result
#cluster errors
?lm_robust()


```


































+ lm_lin  
  *used when independent variable undergoes treatment; includes pre treatment co-variates and then passed through lm_robust
  *
```{r}
?lm_lin()
#lm_lin(data=mroz, hours~kids, covariates=age)
summary(mroz)

#create a dummy for having kids
 
```

+ iv_robust 
  *2SLS estimation of IV regression
```{r}

# good IV = correlated with x but not with u; maybe experience 

iv_robust(data=wage2, wage~educ|IQ)



```
+ difference_in_means -
```{r}
#for variety of designs - unit randomized, cluster randomized, block randomized, matched pairs, etc
#effects estimated with condition 1 = control and condition 2 = treatment 


```
+ horvitz_thompson - 
```{r}
#population-level means and reg coeff 
install.packages("survey")
library(survey)

View 
```


## Screenreg

```{r}

#LEARN
```

 




