---
title: "Guide to estimatr"
author: "Aakriti"
date: '2023-04-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Guide To Estimatr

```{r}
install.packages("estimatr")
library(estimatr)


```

## What are robust standard errors?
Robust standard errors are a way to get unbiased standard errors under heteroskedasticity.Heteroskedasticity --> Gauss MArkov violated --> estimates are not BLUE
#https://economictheoryblog.com/2016/08/07/robust-standard-errors

Economists care because they want to know how accurate their sample is compared to the population.Error more --> mean of sample more spread out --> representation less accurate 
#https://s4be.cochrane.org/blog/2018/09/26/a-beginners-guide-to-standard-deviation-and-standard-error

```{r}
install.packages("wooldridge")
library(wooldridge)
install.packages("npsf")
library(npsf)
data(mroz)
View(mroz)
?mroz

data(hprice1)
View(hrpice)
data("nswre74")

View(wage1)

```


```{r}
reg <- lm(data=wage1, wage~educ+exper)
summary(reg)


```
https://declaredesign.org/r/estimatr/articles/mathematical-notes.html for the explanations



## Estimators provided by Estimatr


+ lm_robust  
  *linear regression models
  *parameters and defaults
  *explanation of results

```{r}
lm_robust(data=wage1, wage~educ+exper) #se_type = classical gives the lm result
?lm_robust()


```

+ lm_lin  
  *used when independent variable undergoes treatment; includes pre treatment co-variates and then passed through lm_robust
  *
```{r}
?lm_lin()
#lm_lin(data=mroz, hours~kids, covariates=age)
summary(mroz)

#create a dummy for having kids
 
```

+ iv_robust 
  *2SLS estimation of IV regression
```{r}

# good IV = correlated with x but not with u; maybe experience 

iv_robust(data=wage2, wage~educ|IQ)



```
+ difference_in_means -
```{r}
#for variety of designs - unit randomized, cluster randomized, block randomized, matched pairs, etc
#effects estimated with condition 1 = control and condition 2 = treatment 


```
+ horvitz_thompson - 
```{r}
#population-level means and reg coeff 
install.packages("survey")
library(survey)

View 
```


## Screenreg

```{r}

#LEARN
```

 




